---
title: "Working with TiffShape Classes"
author: "TiffPlotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with TiffShape Classes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

## Overview

This vignette demonstrates how to:

1. Create each concrete `TiffShape` class.
2. Shift and resize shapes with the `shape_*` generics.
3. Use `ggplot2` to visualize each manipulation.

```{r libraries}
library(TiffPlotR)
library(ggplot2)
library(tidyverse)
```

## A plotting helper

We will draw all examples on the same coordinate system so shape changes are easy to compare.

```{r helper-plot}
empty_canvas <- function(xlim = c(0, 100), ylim = c(0, 100)) {
  ggplot() +
    coord_fixed(xlim = xlim, ylim = ylim) +
    theme_minimal(base_size = 12) +
    labs(x = "x", y = "y")
}
```

## Create all TiffShape classes

### TiffRect

```{r create-rect}
rect0 <- TiffRect(
  xmin = 15,
  xmax = 45,
  ymin = 20,
  ymax = 60,
  name = "rect0"
)

rect0
```

```{r plot-rect}
empty_canvas() %>%
  shape_annotate(rect0, color = "#1f77b4", fill = "#1f77b4", alpha = 0.25) +
  ggtitle("TiffRect")
```

### TiffEllipse

```{r create-ellipse}
ellipse0 <- TiffEllipse(
  x0 = 65,
  y0 = 40,
  radius_x = 18,
  radius_y = 10,
  name = "ellipse0"
)

ellipse0
```

```{r plot-ellipse}
empty_canvas() %>%
  shape_annotate(ellipse0, color = "#2ca02c", fill = "#2ca02c", alpha = 0.25) +
  ggtitle("TiffEllipse")
```

### TiffPolygon

```{r create-polygon}
poly0 <- TiffPolygon(
  x = c(20, 35, 50, 30),
  y = c(75, 92, 78, 65),
  name = "poly0"
)

poly0
```

```{r plot-polygon}
empty_canvas() %>%
  shape_annotate(poly0, color = "#d62728", fill = "#d62728", alpha = 0.25) +
  ggtitle("TiffPolygon")
```

## Shift shapes with shape_shift

```{r shift-shapes}
rect_shifted <- shape_shift(rect0, dx = 12, dy = -8)
ellipse_shifted <- shape_shift(ellipse0, dx = -10, dy = 15)
poly_shifted <- shape_shift(poly0, dx = 25, dy = -12)
```

```{r plot-shift-rect}
empty_canvas() %>%
  shape_annotate(rect0, color = "grey50", fill = "grey70", alpha = 0.2) %>%
  shape_annotate(rect_shifted, color = "#1f77b4", fill = "#1f77b4", alpha = 0.3) +
  ggtitle("shape_shift on TiffRect (gray = original, blue = shifted)")
```

```{r plot-shift-ellipse}
empty_canvas() %>%
  shape_annotate(ellipse0, color = "grey50", fill = "grey70", alpha = 0.2) %>%
  shape_annotate(ellipse_shifted, color = "#2ca02c", fill = "#2ca02c", alpha = 0.3) +
  ggtitle("shape_shift on TiffEllipse (gray = original, green = shifted)")
```

```{r plot-shift-polygon}
empty_canvas() %>%
  shape_annotate(poly0, color = "grey50", fill = "grey70", alpha = 0.2) %>%
  shape_annotate(poly_shifted, color = "#d62728", fill = "#d62728", alpha = 0.3) +
  ggtitle("shape_shift on TiffPolygon (gray = original, red = shifted)")
```

## Resize shapes with shape_resize_abs

```{r resize-abs}
rect_abs <- shape_resize_abs(rect0, width = 18, height = 18, anchor = "center")
ellipse_abs <- shape_resize_abs(ellipse0, width = 20, height = 30, anchor = "center")
poly_abs <- shape_resize_abs(poly0, width = 35, height = 20, anchor = "center")
```

```{r plot-resize-abs}
empty_canvas() %>%
  shape_annotate(rect0, color = "grey50", fill = "grey70", alpha = 0.15) %>%
  shape_annotate(rect_abs, color = "#1f77b4", fill = "#1f77b4", alpha = 0.25) %>%
  shape_annotate(ellipse0, color = "grey50", fill = "grey70", alpha = 0.15)  %>%
  shape_annotate(ellipse_abs, color = "#2ca02c", fill = "#2ca02c", alpha = 0.25)  %>%
  shape_annotate(poly0, color = "grey50", fill = "grey70", alpha = 0.15)  %>%
  shape_annotate(poly_abs, color = "#d62728", fill = "#d62728", alpha = 0.25) +
  ggtitle("shape_resize_abs (gray = original, color = resized)")
```

## Resize shapes with shape_resize_mult

```{r resize-mult}
rect_mult <- shape_resize_mult(rect0, fx = 1.5, fy = 0.7)
ellipse_mult <- shape_resize_mult(ellipse0, fx = 0.6, fy = 1.4)
poly_mult <- shape_resize_mult(poly0, fx = 1.8, fy = 1.2)
```

```{r plot-resize-mult}
empty_canvas() %>%
  shape_annotate(rect0, color = "grey50", fill = "grey70", alpha = 0.15) %>%
  shape_annotate(rect_mult, color = "#1f77b4", fill = "#1f77b4", alpha = 0.25) %>%
  shape_annotate(ellipse0, color = "grey50", fill = "grey70", alpha = 0.15) %>%
  shape_annotate(ellipse_mult, color = "#2ca02c", fill = "#2ca02c", alpha = 0.25) %>%
  shape_annotate(poly0, color = "grey50", fill = "grey70", alpha = 0.15) %>%
  shape_annotate(poly_mult, color = "#d62728", fill = "#d62728", alpha = 0.25) +
  ggtitle("shape_resize_mult (gray = original, color = resized)")
```

## Annotate centers

```{r centers}
empty_canvas() %>%
  shape_annotate(rect0, color = "#1f77b4", fill = NA, alpha = 0.6) %>%
  shape_annotate(ellipse0, color = "#2ca02c", fill = NA, alpha = 0.6) %>%
  shape_annotate(poly0, color = "#d62728", fill = NA, alpha = 0.6) %>%
  shape_annotate(rect0, annotate_center = TRUE, color = "#1f77b4", alpha = 1, size = 3) %>%
  shape_annotate(ellipse0, annotate_center = TRUE, color = "#2ca02c", alpha = 1, size = 3) %>%
  shape_annotate(poly0, annotate_center = TRUE, color = "#d62728", alpha = 1, size = 3) +
  ggtitle("shape_annotate(..., annotate_center = TRUE)")
```

## Working with ggplot2

You can annotate shapes in three complementary styles.

### 1) magrittr pipe with shape_annotate

```{r ggplot-pipe-style}
empty_canvas() %>%
  shape_annotate(rect0, color = "#1f77b4", fill = "#1f77b4", alpha = 0.2) %>%
  shape_annotate(ellipse0, color = "#2ca02c", fill = NA, alpha = 0.8) %>%
  shape_annotate(poly0, color = "#d62728", fill = "#d62728", alpha = 0.15) +
  ggtitle("Pipe style: shape_annotate")
```

### 2) Native ggplot2 `+` with shape objects

When using `+` directly with shape objects, default annotation aesthetics are used.

```{r ggplot-plus-shape}
empty_canvas() +
  rect0 +
  ellipse0 +
  poly0 +
  ggtitle("Plus style: ggplot + shape")
```

### 3) Native ggplot2 `+` with shape_layer for custom styling

Use `shape_layer()` when you want `+` chaining plus explicit annotation parameters.

```{r ggplot-plus-shape-layer}
empty_canvas() +
  shape_layer(rect0, color = "#1f77b4", fill = "#1f77b4", alpha = 0.2) +
  shape_layer(ellipse0, color = "#2ca02c", fill = NA, alpha = 0.9, linewidth = 1.1) +
  shape_layer(poly0, color = "#d62728", fill = "#d62728", alpha = 0.15) +
  shape_layer(poly0, annotate_center = TRUE, color = "black", size = 2.5, alpha = 1) +
  ggtitle("Plus style with custom aesthetics: shape_layer")
```

## Multi-row shapes

All three shape classes support storing multiple rows in one object.

```{r multi-row-create}
rect_multi <- TiffRect(
  xmin = c(5, 20, 35),
  xmax = c(15, 30, 50),
  ymin = c(5, 10, 20),
  ymax = c(18, 24, 34),
  name = c("r", "r", "r")
)

ellipse_multi <- TiffEllipse(
  x0 = c(70, 82, 90),
  y0 = c(70, 82, 65),
  radius_x = c(6, 4, 5),
  radius_y = c(3, 7, 4),
  name = c("e", "e", "e")
)

poly_multi <- TiffPolygon(
  x = list(
    c(55, 62, 58),
    c(60, 68, 70, 64),
    c(48, 54, 50)
  ),
  y = list(
    c(20, 22, 30),
    c(32, 30, 38, 42),
    c(40, 46, 52)
  ),
  name = c("p", "p", "p")
)

rect_multi
ellipse_multi
poly_multi
```

```{r multi-row-plot-base}
empty_canvas() %>%
  shape_annotate(rect_multi, color = "#1f77b4", fill = "#1f77b4", alpha = 0.2) %>%
  shape_annotate(ellipse_multi, color = "#2ca02c", fill = "#2ca02c", alpha = 0.2) %>%
  shape_annotate(poly_multi, color = "#d62728", fill = "#d62728", alpha = 0.2) +
  ggtitle("Multi-row TiffShape objects in one plot")
```

```{r multi-row-shift}
rect_multi_shift <- shape_shift(rect_multi, dx = 8, dy = -3)
ellipse_multi_shift <- shape_shift(ellipse_multi, dx = -6, dy = 5)
poly_multi_shift <- shape_shift(poly_multi, dx = 4, dy = -6)

empty_canvas() %>%
  shape_annotate(rect_multi, color = "grey50", fill = "grey70", alpha = 0.12) %>%
  shape_annotate(ellipse_multi, color = "grey50", fill = "grey70", alpha = 0.12) %>%
  shape_annotate(poly_multi, color = "grey50", fill = "grey70", alpha = 0.12) %>%
  shape_annotate(rect_multi_shift, color = "#1f77b4", fill = "#1f77b4", alpha = 0.25) %>%
  shape_annotate(ellipse_multi_shift, color = "#2ca02c", fill = "#2ca02c", alpha = 0.25) %>%
  shape_annotate(poly_multi_shift, color = "#d62728", fill = "#d62728", alpha = 0.25) +
  ggtitle("Multi-row shape_shift (gray = original, color = shifted)")


```

```{r multi-row-resize}
rect_multi_resized <- shape_resize_mult(rect_multi, fx = c(1.0, 1.4, 0.7), fy = c(1.2, 0.9, 1.5))
ellipse_multi_resized <- shape_resize_mult(ellipse_multi, fx = c(1.3, 0.7, 1.1), fy = c(0.8, 1.4, 1.2))
poly_multi_resized <- shape_resize_mult(poly_multi, fx = c(0.8, 1.5, 1.2), fy = c(1.4, 0.9, 0.8))

empty_canvas() %>%
  shape_annotate(rect_multi, color = "grey50", fill = "grey70", alpha = 0.12) %>%
  shape_annotate(ellipse_multi, color = "grey50", fill = "grey70", alpha = 0.12) %>%
  shape_annotate(poly_multi, color = "grey50", fill = "grey70", alpha = 0.12) %>%
  shape_annotate(rect_multi_resized, color = "#1f77b4", fill = "#1f77b4", alpha = 0.25) %>%
  shape_annotate(ellipse_multi_resized, color = "#2ca02c", fill = "#2ca02c", alpha = 0.25) %>%
  shape_annotate(poly_multi_resized, color = "#d62728", fill = "#d62728", alpha = 0.25) +
  ggtitle("Multi-row shape_resize_mult with per-row scale factors")
```

## Backward compatibility

Rectangle-specific helpers remain available and dispatch the new shape generics.

```{r legacy-rect}
legacy_rect <- rect_shift(rect0, dx = 5, dy = 5)
identical_coords <- identical(legacy_rect@coords, shape_shift(rect0, dx = 5, dy = 5)@coords)
identical_coords
```
